---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# statable -  Run Stata from R

<!-- badges: start -->
<!-- badges: end -->

statable is an R interface to Stata that allows you to combine the strengths of both R and Stata into one seamless workflow.
You can prepare and manipulate your data in R, then quickly switch to Stata to perform specialised analyses, all without leaving your R environment.
This integration streamlines your work, saves time and reduces the need to manually transfer data between software.

statable provides the following functionality:

* Run Stata commands from R
* Transfer datasets between R and Stata
* Run Stata "chunks" in knitr or R Markdown documents
* Work with multiple Stata sessions at the same time

## Installation

`statable` is not currently on CRAN.
You can install the development version from [GitHub](https://github.com/james-atkins/statable) with:

``` r
# install.packages("pak")
pak::pak("james-atkins/statable")
```

## Usage

For more information, please read the [documentation](https://statable.jamesatkins.com).

```{r usage}
# First, load the package. All statable commands start with `stata_`.
library(statable)

# statable should find Stata automatically but, if necessary, you can
# manually set the path to the Stata executable
# stata_path("C:/Program Files/Stata18/StataSE-64.exe")

# `stata_run()` runs Stata commands in the same session.
# All your data and results stay in one place, allowing you to easily build on
# your previous commands.
stata_run("sysuse auto")
stata_run(c("summarize", "regress price length weight"))

# To pass data from R to Stata, you can use the function `stata_data_in()`
data("mtcars")
stata_data_in(mtcars, clear = TRUE)
stata_run("describe")

# Likewise, `stata_data_out()` returns the current Stata dataset into R
stata_run("rename mpg miles_per_gallon")
mtcars_stata <- stata_data_out()
mean(mtcars_stata$miles_per_gallon)

# You can also reference R data frames using Stata globals of the form
# `$R_dataframe_name`.
# This allows you to use data in R with a variety of Stata commands, such as
# `merge`, `joinby` or `append`.
stata_run("use $R_mtcars, clear")
stata_run("summarize")
stata_run("append using $R_mtcars")
stata_run("display _N")
```

## knitr

`statable` integrates with knitr allowing you to run Stata code chunks in knitr or rmarkdown documents.
To do so, you must first import the package with `library(statable)` so knitr knows what to do with `stata` chunks.
Then add a chunk like the following.

````{verbatim}
```{stata}
use $R_mtcars, clear
summarize
```
````
