---
title: "Introduction to statable"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to statable}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The aim of statable is to enable you to combine the strengths of both R and Stata into one seamless workflow.
This document introduces you to statable’s basic functionality.

To start using statable, you must first load the package.

```{r setup}
library(statable)
```

<i class="fa fa-info-circle" style="color: #007bff;"></i> **Note: ** All statable functions start with the prefix `stata_`.

# Finding Stata

In order to run, statable needs the path to your Stata executable.
It should find this path automatically, but if it cannot you can specify it using `stata_path()`.
This function may also be useful if you have multiple versions of Stata installed and want to run an older version.

```{r stata_find, eval=FALSE}
# For example on Windows
stata_path("C:/Program Files/Stata18/StataSE-64.exe")

# or on Mac
stata_path("/Applications/Stata/StataSE.app/Contents/MacOS/stata-se")

# or on Linux
stata_path("/opt/stata18/stata-mp")
```

# Running Commands

To run a Stata command from R, simply use the `stata_run()` function.
You can pass a single Stata command to this function or a vector of commands.

```{r stata_run}
stata_run("sysuse auto")
stata_run(c("summarize", "regress price length weight"))
```

See how the second set of commands used the `auto` dataset that we loaded with the first command.
This works because, by default, `stata_run()` uses the same Stata session every time you run a command.
This means all your data and results stay in one place, allowing you to easily build on your previous commands.
If needed, you can also run multiple Stata sessions at the same time but most users can use statable without worrying about session management.

# Transferring Data

## Functions

statable makes it is easy to transfer datasets between R and Stata.
This is useful when you want to use Stata’s powerful statistical capabilities on data that you initially cleaned and tidied in R.

To pass data from R to Stata, use the function `stata_data_in()`.
This function takes an R data frame and sends it to the Stata session that you are working with.
Behind the scenes, `stata_data_in()` exports the R data frame to a `.dta` file using the `haven::write_dta` function.

Similarly, if you have been working with a dataset in Stata and want to bring that data back into R, use the function `stata_data_out()`.
This function retrieves the current dataset from the active Stata session and returns it as a data frame.

```{r data}
data("mtcars")

stata_data_in(mtcars, clear = TRUE)
stata_run("summarize")

stata_run("rename mpg miles_per_gallon")
mtcars_stata <- stata_data_out()

mean(mtcars_stata$miles_per_gallon)
```

## Magic Macros

In addition to these direct data transfer functions, statable provides a more general way to work with data between R and Stata using global macros.
In Stata, global macros of the form `$R_dataframe_name` are used to reference R data frames.
When you create a global macro in this format, statable automatically exports the specified R data frame to a `.dta` file, which is the file format used by Stata for datasets.
The value of the global macro is then set to the path of this `.dta` file.

This functionality allows you to seamlessly use R data frames with a variety of Stata commands that require a path to a dataset, such as `merge`, `joinby`, or `append`.
For example, in the following code, `$R_mtcars` references the R data frame called `mtcars` in Stata, allowing you to `use` it and then `append` it.

```{r macros}
# statable sets the gloal R_mtcars to be equal to the path of a dta file
# containing the data frame mtcars.
stata_run("use $R_mtcars, clear")
stata_run("summarize")
stata_run("display _N")

stata_run("append using $R_mtcars")
stata_run("display _N")
```

<i class="fa fa-info-circle" style="color: #007bff;"></i> **Note: ** R allows both variable names and column names to have dots in them but this is **not** allowed by Stata. You will have to rename them first.
